---
title: 'New Haven GI CATE'
author: 'Dexter H. Locke, PhD'
date: '`r format(Sys.time())`'
output: html_document
editor_options:
  chunk_output_type: console
---

Questions? 
dexter.locke@gmail.com; dexter.locke@usda.gov; Twitter: @DextraordinaryH; web: http://dexterlocke.com/


This RMarkdown is part of the larger package:

Locke, D. H. (2021). Green stormwater infrastructure in New Haven, CT and 
socioeconomic variables within neighborhoods and census block groups. Fort Collins, CO: Forest Service Research Data Archive. https://doi.org/10.2737/RDS-2021-0104

which contains data and this code to replicate some of the analyses contained in 

Locke, D.H., Phillips de Lucas, A.K., Murphy-Dunning, C., Henning, D., Zinn, G. 2021. Contextual Considerations of Green Infrastructure Siting. Cities And The Environment. 10.15365/cate.2021.140202 https://digitalcommons.lmu.edu/cate/vol14/iss2/2/

Attribution is appreciated. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 0 set up, get packages - if needed - and load them
```{r}
`%nin%` <- Negate(`%in%`) # custom function

# where are we
getwd()
list.files()

# load useful packages
packs <- c('tidyverse',           # nearly a must-have
           'tidylog',             # makes tidyverse surly (very verbose)
           'magrittr',            # ALL of the pipes, not just %>%
           'janitor',             # data cleaning and pipe-friendly cross tabulations
           'santoku',             # chop! (or Kiru)
           'Hmisc',               # for rcorr
           'sjPlot',              # for sjt.corr
           'sf',                  # Simple Features, modern spatial data in R
           'spdep',               # SPatial DEPendence - a juggernaut
           'spatialreg',          # might not use, but spatial regression, split from spdep
           'mapview',             # quick webmaps for zooming/ panning around
           'rgdal',
           'gdalUtils',
           'raster',
           'tmap',
           'epitools',            # contains pois.exact function for 95% CI's around standardized rates
           'tidycensus')          # access to Census data in a tidy way
                                  # additional information and examples
                        # https://walkerke.github.io/tidycensus/articles/basic-usage.html#searching-for-variables
                        # https://walkerke.github.io/tidycensus/articles/spatial-data.html)            # 

# check for all of the libraries
if(length(setdiff(packs, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packs, rownames(installed.packages())))
}

# load them all at once
vapply(packs, library, character.only = TRUE, logical(1), logical.return = TRUE, quietly = TRUE)


```


## 1 read in the data
### A neighborhoods
```{r}
# ESRI likes to rename columns, these are the column names that are more manageable and readable
neigh_shp_names <- c(
    'rowname.x'
  , 'id'
  , 'STATE'
  , 'COUNTY'
  , 'CITY'                               
  , 'Neighborhood'                       
  , 'REGIONID.x'                         
  , 'Total population_2017'              
  , 'Population under age 18_2017'       
  , 'Percent under age 18_2017'          
  , 'Population ages 65+_2017'           
  , 'Percent ages 65+_2017'              
  , 'Latino population_2017'             
  , 'Percent Latino_2017'                
  , 'White population_2017'              
  , 'Percent white_2017'                 
  , 'Black population_2017'              
  , 'Percent black_2017'                 
  , 'Other race population_2017'         
  , 'Percent other race_2017'            
  , 'Foreign-born population_2017'       
  , 'Percent foreign-born_2017'          
  , 'Total households_2017'              
  , 'Owner-occupied households_2017'     
  , 'Homeownership rate_2017'            
  , 'Cost-burdened households_2017'      
  , 'Cost-burden rate_2017'              
  , 'Poverty status known_2017'          
  , 'Population in poverty_2017'         
  , 'Poverty rate_2017'                  
  , 'Low-income population_2017'         
  , 'Low-income rate_2017'               
  , 'Under 18, poverty status known_2017'
  , 'Children in poverty_2017'           
  , 'Child poverty rate_2017'            
  , 'Low-income children_2017'           
  , 'Child low-income rate_2017'         
  , 'Ages 65+, poverty status known_2017'
  , 'Seniors in poverty_2017'            
  , 'Senior poverty rate_2017'           
  , 'Low-income seniors_2017'            
  , 'Senior low-income rate_2017'        
  , 'Life expectancy_2010-2015'          
  , 'High blood pressure_2015'           
  , 'Coronary heart disease_2016'        
  , 'Current asthma_2016'                
  , 'Diabetes_2016'                      
  , 'Annual checkup_2016'                
  , 'Dental visit_2016'                  
  , 'Health insurance_2016'              
  , 'Current smoking_2016'               
  , 'Sleep <7 hours_2016'                
  , 'gi_counts'                          
  , 'a_tree'                             
  , 'a_grass'                            
  , 'a_soil'                             
  , 'a_water'                            
  , 'a_building'                         
  , 'a_road'                             
  , 'a_other_paved'                      
  , 'a_total'                            
  , 'p_tree'                             
  , 'p_grass'                            
  , 'p_soil'                             
  , 'p_water'                            
  , 'p_building'                         
  , 'p_road'                             
  , 'p_other_paved'                      
  , 'Impervious Surface (%)'             
  , 'REGIONID.y'                         
  , 'gi_per_cap'                         
  , 'gi_per_1000'                        
  , 'expected'                           
  , 'std_rate'                           
  , 'rowname.y'                          
  , 'x'                                  
  , 'pt'                                 
  , 'rate'                               
  , 'lower'                              
  , 'upper'                              
  , 'conf.level'
  , 'geometry'                           
  )

# read in the data
neigh_sf <- st_read(paste0(getwd(), '/input_data/neighborhoods.shp')) %>% 
  rename_at(vars(everything()), ~neigh_shp_names)

```

### B GSI and watersheds
```{r}

gi <- st_read(paste0(getwd(), '/input_data/gi_cartographic.shp')) %>% 
  mutate(label = 'Green Stormwater Infrastructure (GSI)')

shed_outlines <- st_read(paste0(getwd(), '/input_data/shed_outlines.shp')) %>% 
  mutate(label = 'Combined Sewer Areas')

dt_shed <- st_read(paste0(getwd(), '/input_data/dt_shed.shp')) %>% 
  mutate(label = 'Downtown Watershed')

```

### C Census block groups
```{r}
# ESRI likes to rename columns, these are the column names that are more manageable and readable
cbg_shape_names <- c(
    'rowname'
  , 'GEOID'
  , 'tot_pop'
  , 'medincomeE'
  , 'pct_white'
  , 'pct_black'             
  , 'pct_am_ind'             
  , 'pct_asian'              
  , 'pct_pac_is'            
  , 'pct_other_r'
  , 'pct_own_occ'
  , 'pct_vac'
  , 'pct_edu'
  , 'gi_counts'
  , 'id'
  , 'a_tree'
  , 'a_grass'
  , 'a_soil'
  , 'a_water'
  , 'a_building'
  , 'a_road'
  , 'a_other_paved'
  , 'a_total'
  , 'p_tree'
  , 'p_grass'
  , 'p_soil'
  , 'p_water'
  , 'p_building'
  , 'p_road'
  , 'p_other_paved'
  , 'Impervious Surface (%)'
  , 'gi_per_cap'
  , 'gi_per_1000'
  , 'expected'
  , 'std_rate'  
  , 'geometry'
  )


nh_cbg_sf <- st_read(paste0(getwd(), '/input_data/census_block_groups.shp')) %>% 
  rename_at(vars(everything()), ~cbg_shape_names)

```

### D CSOs
```{r}

cso <- st_read(paste0(getwd(), '/input_data/CSOs.shp')) %>% 
  mutate(label = 'Combined Sewer\nOverflows (CSOs)')

```


## 2 Data outputs
### A Figure 1: study area - poverty, CSOs, downtown watershed
```{r}

# prep
nh_cbg_sf %>% group_by() %>% summarise() -> outline # city outline, for cartography (not used)

neigh_sf %<>% # make nicer labels for the map
  mutate(labs = as.factor(Neighborhood),
         labs = recode(labs, 
                       'West Rock' = 'West\nRock',
                       'West River' = 'West\nRiver',
                       'Fair Haven Heights' = 'Fair Haven\nHeights',
                       'East Shore' = 'East\nShore',
                       'Long Wharf' = 'Long\nWharf',
                       'Wooster Square' = 'Wooster\nSquare',
                       'Beaver Hills' = 'Beaver\nHills',
                       'Prospect Hill' = 'Prospect\nHill'))


neigh_text_diff <- c('Dwight', 'West River', 'Hill', 'Long Wharf', 'West Rock')

neigh_sf %>% filter(Neighborhood %in% neigh_text_diff) -> black_label_neighs

neigh_sf %>% filter(Neighborhood %nin% neigh_text_diff) -> white_label_neighs


# Revised Figure 1
# neighborhoods by poverty and with CSOs
# begin map
tm_shape(neigh_sf) +
  tm_polygons('Poverty rate_2017', title = 'Poverty Rate, Year 2017', legend.reverse = TRUE,
              palette = 'viridis') + # palette takes the RColorBrewer values                                                               # RColorBrewer::display.brewer.all()
tm_shape(neigh_sf) +
  tm_borders('white') +
tm_shape(dt_shed) +
  tm_polygons('label', title = '', alpha = 0, lwd = 3) +
tm_shape(white_label_neighs) +
  tm_text('labs', size = .6, col = 'white', shadow = TRUE) +
tm_shape(black_label_neighs) +
  tm_text('labs', size = .6, col = 'black', shadow = TRUE) +
tm_shape(cso) +
  tm_dots('label', title = '', size = .2, col = 'black', shape = 21) +
tm_add_legend(type = 'symbol', labels = 'Combined Sewer Overflows (CSOs)',
              col = 'black', shape = 21, size = 1.1) +
  tm_layout(  legend.text.size = 1.1
            , legend.title.size = 1.5
            , legend.width = 1.5
            )
  # tm_credits(text = paste0('created on ', gsub('[[:punct:]]', '_', Sys.time()), ' by DHL'),
  #            position = c('left', 'bottom')) + 
  # tm_credits(text = 'A', size = 1.75,
  #            position = c('right', 'top'))

tmap_save(filename = paste0('output_data/graphics/Fig_1_poverty_CSO_',
                            gsub('[[:punct:]]', '_', Sys.time()), '.png'),
          width = 3.25*2,
          height =3.25*2,
          units = 'in')
```

### B Figure 2: CSO areas, GSI locations and the downtown watershed
```{r}

# Revised Figure 2
tm_shape(shed_outlines) + 
  tm_polygons() + 
tm_shape(neigh_sf) + 
  tm_polygons('Poverty rate_2017', title = 'Poverty Rate,\nYear 2017', legend.reverse = TRUE,  
              palette = 'viridis') + 
tm_shape(shed_outlines) +
  tm_polygons('label', alpha = .3, title = '', lwd = 2, border.col = 'black') +
tm_shape(dt_shed) +
  tm_polygons('label', title = '', alpha = 0, lwd = 3, border.col = 'red') +
# tm_shape(dt_shed) + 
#   tm_borders('red', lwd = 2) +
tm_shape(gi) +
  tm_dots('label', size = .05, col = 'black', shape = 3) +
  tm_add_legend(type = 'symbol', labels = 'Green Stormwater Infrastructure (GSI)',
              border.col = 'black', col = 'black', shape = 3, size = .75)

tmap_save(filename = paste0('output_data/graphics/Fig_2_CSO_GSI_',
                            gsub('[[:punct:]]', '_', Sys.time()), '.png'),
          width = 3.25*2,
          height =3.25*2,
          units = 'in')


```

### C Figure 3: Neighborhood-level Odds Ratios
```{r}

# neighborhood odds ratios
neigh_sf %>%
  mutate(Neighborhood = ifelse(Neighborhood == 'Hill', 'the Hill', Neighborhood)) %>% 
  dplyr::filter(Neighborhood != 'Long Wharf') %>% 
  ggplot(aes(x = rate, y = reorder(Neighborhood, `Poverty rate_2017`), col = `Impervious Surface (%)`)) + 
  scale_x_continuous(name = 'Odds Ratio', breaks = seq(0, 9)) + 
  geom_vline(xintercept = 0) + 
  geom_vline(xintercept = 1, col = 'gray', size = 2) + 
  # geom_vline(xintercept = 2, lty = 2) + # to denote really extreme values
  geom_errorbarh(aes(xmin = lower, xmax = upper), col = 'gray') + 
  geom_point(size = 3) + 
  geom_point(size = 3, shape = 21, color = 'black') + 
  viridis::scale_color_viridis(option = 'inferno', direction = -1) + 
  labs(title = 'Green Stormwater Infrastructure by Neighborhood',
       subtitle = 'New Haven, CT',
       y = 'Neighborhood') + 
  geom_segment(x = 5.5,
               y = 15,
               color = 'black',
               xend = 5.5,
               yend = 5,
               arrow = arrow(length = unit(0.03, "npc"), ends = 'both')) +
  annotate('text',
           x = 6,
           y = 14,
           size = 4,
           hjust = 0,
           label = 'Higher Poverty Rate') +
  annotate('text',
           x = 6,
           y = 7,
           size = 4,
           hjust = 0,
           label = 'Lower Poverty Rate,\nYear 2017') +
  theme_classic(12) + 
  theme(legend.position = 'bottom', plot.title = element_text(hjust = 0))

# save out
ggsave(filename = paste0('output_data/graphics/Fig_3_odds_ratio_',
                            gsub('[[:punct:]]', '_', Sys.time()), '.png'),
          width = 7,
          height =6,
          units = 'in')

```

### D Table 1: Neighborhood Descriptive Statistics
```{r}


neigh_sf %>% 
  st_drop_geometry() %>% 
  mutate('Latino Population (%)' = `Percent Latino_2017`*100,
         'White Population (%)' = `Percent white_2017`*100,
         'African American (%)' = `Percent black_2017`*100,
         'Foreign Born (%)' = `Percent foreign-born_2017`*100,
         'Poverty Rate' = round(`Poverty rate_2017`, 2),
         'Poverty Rate' = round(`Poverty rate_2017`, 6),
         'GSI* Odds Ratio' = round(rate, 2),
         'lower 95% CI' = round(lower, 2),
         'upper 95% CI' = round(upper, 2),
         'GSI* per 1,000 residents' = round(gi_per_1000, 3),
         Neighborhood = ifelse(Neighborhood == 'Hill', 'the Hill', Neighborhood),
         `Impervious Surface (%)` =  round(`Impervious Surface (%)`)) %>% 
  tidylog::select(Neighborhood,
                  `Impervious Surface (%)`,
         'Total Population (2017)' = `Total population_2017`,
         'Latino Population (%)',
         'White Population (%)',
         'African American (%)',
         'Foreign Born (%)',
         'Poverty Rate',
         'GSI* (n)' = gi_counts,
         #'GSI* per capita' = gi_per_cap,
         'GSI* per 1,000 residents',
         'GSI* Odds Ratio',
         'lower 95% CI',
         'upper 95% CI') %>% 
  write.csv(., file = paste0(getwd(), '/output_data/tables/Table_1_neigh_summ_stats_slim_', Sys.Date(), '.csv'),
            row.names = F)

```


### E Table 2: Census Block Group Correlations
```{r}

# pretty up names in the selection for the correlation matrix
nh_cbg_sf %>% 
  st_drop_geometry() %>% 
  tidylog::select(#GEOID,
          'GSI* (n)' = gi_counts
         , 'GSI* per capita' = gi_per_cap
         , 'GSI* Odds Ratio' = std_rate
         
         , 'Median Household Income ($)' = medincomeE
         , 'White Population (%)' = pct_white
         , 'African American (%)' = pct_black
         , 'Owner Occupied (%)' = pct_own_occ
         , 'Vacant Housing (%)' = pct_vac
         , 'Educational Attainment**' = pct_edu) %>% 
  tab_corr(.
           ,
           corr.method = 'spearman',
         triangle = 'lower',
         p.numeric = TRUE, # remove the "#" below to save the file
         file = paste0(getwd(), '/output_data/tables/Table_2_cbg_demo_corrs_', Sys.Date(), '.html')
         )


```



## CITE your sources!!!
```{r}
lapply(packs, citation); rm(packs)
sessionInfo()
```

Last Knit on `r format(Sys.time())`

# end
