---
title: "New_Haven_street_trees_redlining"
author: "Dexter H. Locke, PhD"
date: "`r format(Sys.time())`"
output: html_document
editor_options: 
  chunk_output_type: console
---

Add in Park Score

Focus on: Wooster square, Beaverponds, Kimberly Field
  and their neighborhoods
    what are their HOLC grades?
    current Census data


Ash Map 3% ash city wide.. but totally clustered.
  Map of single species?
  Alpha vs Beta diversity?
  

TODO
  graphs to make
      Three inventoried parks relative to all street trees previously in the inventory (exclude data previously collected)
  When trees were planted? (inv_date, plntdate)
  bin inspect_dt (data)
  street trees per length
  group dbh into classes

species accumulation curves




MODs
GEOFACET (DBH not ideal)
  species
  species diversity
  stems
  stems per length
  ... something that shows inequality. Where can problems be addressed?




DONE powerpoint slides
  BIKE map of example track
  estimates next to Carly's

  unit of analysis matches unit of management
  
  add notes, next steps

  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# 0 load libraries and read in data----
```{r message=FALSE}
# Load libraries #### from Carly 
packs <-c('tidyverse'   # cuz
          , 'tidylog'   # prints out what was done in dplyr and tidr
          , 'janitor'   # cleans things up, also pipe-friendly cross-tabulations
          , 'sf'        # Simple Features, for spatial data support
          , 'mapview'   # web maps for zooming and panning around
          , 'nngeo'     # st_nn is faster and more straight forward than st_* alternatives
          , 'parallel'  # parallel processing, vroom, vroom
          , 'ggrepel'   # to move text labels on graphs
          , 'vegan'     # vegetation analyses
          ) 
          

if (length(setdiff(packs, rownames(installed.packages()))) > 0) {
install.packages(setdiff(packs, rownames(installed.packages())))
}

# load all packages at once.
vapply(packs, library, character.only = TRUE, logical(1), logical.return = TRUE, quietly = TRUE)


list.files()


mapviewOptions(fgb = FALSE) # fixes mapview (something about new versions.. )


# for reproducibility, we should have the same random draws. setting the seed ensures that is the case.
set.seed(19870630)



holc_pal <- c('#92BC6B' # green
              , '#92C7C9' # blue
              , '#E7DC6B' # yellow
              , '#E47D67' # red
              #, '#A9A9A9'
              ) # dark gray)


```


## i neigh geofacet
```{r}

library(geofacet)
us_state_grid1 # sample, it goes: row, col, code, name


(
  nh_neighs_geofacet <- tribble(
    ~row, ~col, ~code,                     ~name,        
    1,    1,    'Amity',                   'Amity',        
    1,    2,    'West Rock',               'West Rock',      
    1,    3,    'Beaver Hills',            'Beaver Hills',       
    1,    4,    'Newhallville',            'Newhallville',      
    2,    5,    'Prospect Hill',           'Prospect Hill',      
    2,    6,    'East Rock',               'East Rock',
    3,    7,    'Fair Haven',              'Fair Haven',
    2,    7,    'Quinnipiac Meadows',     'Quinnipiac\nMeadows',
    2,    2,    'Westville',               'Westville',
    2,    3,    'Edgewood',                'Edgewood',
    2,    4,    'Dixwell',                 'Dixwell',
    3,    5,    'Downtown',                'Downtown',
    # 3,    6,    'Wooster Square / Mill River',          'Wooster Square / Mill River',
    3,    6,    'Wooster Square',          'Wooster Square\n/ Mill River',
    3,    4,    'Dwight',                  'Dwight',
    3,    3,    'West River',              'West River',
    4,    4,    'Hill',                    'Hill',
    4,    5,    'Long Wharf',              'Long Wharf',
    4,    7,    'Fair Haven Heights',      'Fair Haven\nHeights',
    5,    7,    'Annex',                   'Annex',
    6,    7,    'East Shore',              'East Shore'
    )
  )

grid_preview(nh_neighs_geofacet)


```




# 1 read in data
## A Neighborhoods
```{r}

(neighs <- st_read('input_data/RDS-2021-0104/Data/neighborhoods.shp', as_tibble = TRUE) |> 
   rename(neighborhood = nghbr))

neighs |> mapview(zcol = 'neighborhood')

neighs |> distinct(neighborhood)

neigh_labs <- neighs |> 
  select(neighborhood) |> 
  st_centroid()

# neigh_labs |> mapview()

```


## B HOLC polygons
```{r}
# '../' means 'up one level'
(holc <- st_read('../../../BaltimoreGIS/BHall/HOLC_X_Sites/HOLC_NewHaven/shp/HOLC_NewHaven.shp') |> 
   st_transform(crs = st_crs(neighs)))


holc |> mapview(zcol = 'holc_grade', col.regions = holc_pal) +
  mapview(neighs |> summarise() |> nngeo::st_remove_holes(), alpha.regions = 0, lwd = 2) 

```



## C parks
```{r}

parks <- st_read('input_data/new haven parks/Parks.shp', as_tibble = TRUE) |> 
  rowid_to_column() |> 
  filter(Category == 'Park') |> # drop golf courses and triangles?
  mutate(parks_to_keep = 
           case_when(
             rowid == '115' ~ 1
             , rowid == '125' ~ 1
             , rowid == '69'  ~ 1
             , rowid == '107' ~ 1
             , TRUE ~ 0 
            )
         , ParkName = 
           case_when(
               ParkName == 'Beaver Pond Park (Blackmon)' ~ 'Beaver Pond Park'
             , ParkName == 'Jackie Robinson School Park' ~ 'Beaver Pond Park'
             , ParkName == 'Vietch Field'                ~ 'Beaver Pond Park'
             , ParkName == 'Bowen Field'                 ~ 'Beaver Pond Park'
             , TRUE ~ ParkName
           )
           ) |> 
  st_transform(crs = st_crs(neighs))


parks |> mapview(col.regions = 'light green')

parks |> st_drop_geometry() |> tabyl(Category)

# parks_to_keep <- c('Kimberly Field'
#                    # , 'Beaver Pond Park (Blackmon)'
#                    # , 'Jackie Robinson School Park'
#                    # , 'Vietch Field'
#                    # , 'Bowen Field'
#                    , 'Wooster Square'
#                    )

parks |> 
  # filter(parks_to_keep == 1) |> 
  group_by(ParkName) |> 
  summarise() |> 
  mapview(
    # col.regions = 'light green'
    )


# park_trees <- trees |> #st_drop_geometry() |> # drops spatial information
#     filter(dbhheight == '4.5ft') |>
#     filter(department == 'Parks' & growspace == 'Yard/Park/Open space') 
# 
# # 2x check
# park_trees |> st_drop_geometry() |> tabyl(growspace)
# 
# park_trees |> mapview()
# 
# 
# park_trees |> 
#   
#   mapview() + 
#   mapview(holc, zcol = 'holc_grade', col.regions = holc_pal) + 
#   mapview(parks |> filter(ParkName %in% parks_to_keep), col.regions = 'dark green', lwd = 2)

```



## D trees (via Davey Resources Group)
```{r}

(
  trees <- st_read('input_data/1655469708431_tree_sites/tree_sites.shp', as_tibble = TRUE) |> 
    clean_names() |> 
    filter(!is.na(nbrhood)) |> 
    st_transform(crs = st_crs(neighs)) |> 
    st_join(holc |> select(holc_grade), join = st_intersects) |> # add holc grade
    st_join(parks |> select(`Park Name` = ParkName, parks_to_keep)
            , join = st_intersects) |> # add holc grade
    mutate(neighborhood = ifelse(nbrhood == 'Wooster Square / Mill River'
                                 , 'Wooster Square', nbrhood)
           , dbh_class = cut(dbh, breaks = seq(0, 100, 5))
           # , dbh_class = fct_lump_n(dbh_class, 9, other_level = '> 45 in')
           , dbh_class = fct_lump_n(dbh_class, 7, other_level = '> 35 in')
           )
    )


# 2x check
trees |> 
  sample_n(1000) |> # random sample 1000 points, larger will be slower to load.
  mapview() + 
  mapview(neighs |> summarise() |> nngeo::st_remove_holes(), alpha.regions = 0, lwd = 2) +
  mapview(neighs, alpha.regions = 0)


# test query 
trees |> 
  st_drop_geometry() |> 
  tabyl(inv_date) |> 
  tibble() |> 
  tail()

trees |> 
  filter(inv_date < '2022-06-15')
  # filter(inv_date > '2022-06-15')
  # filter(inv_date == inspect_dt)


```


# 2 EDA
```{r eval=FALSE, include=FALSE}
trees |> glimpse()

trees |> st_drop_geometry() |> tabyl(neighborhood)
trees |> st_drop_geometry() |> tabyl(department)
trees |> st_drop_geometry() |> tabyl(growspace)
trees |> st_drop_geometry() |> tabyl(dbh_class)
# trees |> st_drop_geometry() |> mutate(dbh_test = fct_lump_n(dbh_class, 7)) |> tabyl(dbh_test)

(trees |> st_drop_geometry() |> # TODO filter by street trees only? 
    # FIXME filter out to keep only parks?
    tabyl(department, growspace) |> 
    tibble() |> 
    pivot_longer(-department) -> tree_space_dept)

tree_space_dept |> 
  ggplot(aes(value, department, fill = name)) + 
  geom_col() + 
  theme_bw(16) + 
  scale_x_continuous(expand = c(0, 0), limits = c(0, 27500)) +
  NULL

(mean_dbh <- trees |> 
  st_drop_geometry() |> # drops spatial information
  filter(dbhheight == '4.5ft') |> 
  pull(dbh) |> 
  mean())

trees |> st_drop_geometry() |> # drops spatial information
  filter(dbhheight == '4.5ft') |> 
  ggplot(aes(dbh)) + 
  geom_density() + 
  theme_bw(16) + 
  geom_vline(xintercept = mean_dbh, color = 'red') + 
  # facet_wrap(~neighborhood) +
  geofacet::facet_geo(~neighborhood, grid = nh_neighs_geofacet, label = 'name') + 
  theme(strip.text.x = element_text(size = 10)) +
  NULL



```


## A DBH: street trees
```{r}

# DBH bins
trees |> st_drop_geometry() |> # drops spatial information
  filter(dbhheight == '4.5ft') |> 
  filter(department == 'Street Tree') |> 
  tabyl(dbh_class) |> 
  tibble() |> 
  drop_na() |> 
  ggplot(aes(dbh_class, n)) + 
  geom_col() + 
  theme_bw(16) + 
  labs(
    title = 'New Haven Street Trees by Size'
    , x = 'Diameter at Breast Height (DBH) class'
    , y = 'Count'
  ) +
  NULL

```


## B DBH by neighborhood: street trees
```{r}

# DBH bins
trees |> st_drop_geometry() |> # drops spatial information
  filter(dbhheight == '4.5ft') |> 
  filter(department == 'Street Tree') |> 
  group_by(neighborhood, dbh_class) |> 
  count() |> 
  drop_na(dbh_class) |> 
  ggplot(aes(dbh_class, n)) + 
  geom_col() + 
  theme_bw() + 
  # facet_wrap(~neighborhood) +
  geofacet::facet_geo(~neighborhood, grid = nh_neighs_geofacet, label = 'name') + 
  theme(strip.text.x = element_text(size = 10)
        , axis.text.x = element_text(angle = 90)
        ) +
  labs(
    title = 'New Haven Street Trees by Size and Neighborhood'
    , x = 'Diameter at Breast Height (DBH) class'
    , y = 'Count'
  ) +
  NULL


```


## C street tree species 
```{r}
# find most common species in each neighborhood
(
  trees |> st_drop_geometry() |> # drops spatial information
    filter(dbhheight == '4.5ft') |>
    filter(department == 'Street Tree') |> 
    group_by(neighborhood, species_co) |> # count species per neighborhood
    count() |>
    arrange(neighborhood, desc(n)) |>
    group_by(neighborhood) |>
    slice(1:5) |>
    ungroup() |>
    rename(Species = species_co) -> top_5_species_neighborhood
  )


top_5_species_neighborhood |>
  # ggplot(aes(n, reorder(Species, n))) +
  ggplot(aes(n, Species)) +
  geom_col() + 
  theme_bw() +
  labs(title = 'Top 5 most common street tree species in New Haven, CT per neighborhood'
       # , subtitle = 'by neighbothood'
       ) + 
  # facet_wrap(~neighborhood) +
  geofacet::facet_geo(~neighborhood, grid = nh_neighs_geofacet, label = 'name', scales = 'free_y') +
  theme(strip.text.x = element_text(size = 10)
        , axis.text.x = element_text(angle = 90)
        ) +
  # coord_flip() +
  NULL


# # https://stackoverflow.com/questions/51710863/how-to-order-bars-within-all-facets
# scale_y_reordered <- function(..., sep = "___") {
#   reg <- paste0(sep, ".+$")
#   ggplot2::scale_y_discrete(labels = function(y) gsub(reg, "", y), ...)
# }
# 
# reorder_within <- function(y, by, within, fun = mean, sep = "___", ...) {
#   new_y <- paste(y, within, sep = sep)
#   stats::reorder(new_y, by, FUN = fun)
# }
# 
# 
# # ggplot(ii, aes(reorder_within(sn, nbr, s), nbr)) +
# #      geom_bar(stat = 'identity') +
# #      scale_x_reordered() +
# #      facet_wrap(.~ s, ncol=2,scales = "free_x") + 
# #     theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5,colour='gray50'))
# 
# 
# 
# scale_x_reordered <- function(..., sep = "___") {
#   reg <- paste0(sep, ".+$")
#   ggplot2::scale_x_discrete(labels = function(x) gsub(reg, "", x), ...)
# }
# 
# reorder_within <- function(x, by, within, fun = mean, sep = "___", ...) {
#   new_x <- paste(x, within, sep = sep)
#   stats::reorder(new_x, by, FUN = fun)
# }
# 
# 
# 
# # find most common species in each neighborhood
# (trees |> st_drop_geometry() |> # drops spatial information
#   filter(dbhheight == '4.5ft') |> 
#   group_by(neighborhood, species_co) |> # count species per neighborhood
#   count() |> 
#   arrange(neighborhood, desc(n)) |> 
#   group_by(neighborhood) |> 
#   slice(1:5) |> 
#   ungroup() |>
#   rename(Species = species_co) -> top_5_species_neighborhood)
# 
# 
# top_5_species_neighborhood |> 
#   ggplot(aes(Species, reorder_within(n, Species, neighborhood))) + 
#   geom_bar(stat = 'identity') + 
#   scale_y_reordered() +
#   # scale_x_reordered() +
#   # theme_bw(16) + 
#   # facet_wrap(~neighborhood) +
#   geofacet::facet_geo(~neighborhood, grid = nh_neighs_geofacet, label = 'name', scales = 'free_y') +
#   theme(strip.text.x = element_text(size = 10)) +
#   coord_flip() + 
#   NULL
# 
#   
# trees |> st_drop_geometry() |> # drops spatial information
#   filter(DBHHEIGHT == '4.5ft') |> 
#   group_by(NBRHOOD) |> 
#   summarise(species_frequency = n_distinct(SPECIES_CO))

```


### a RAC test
```{r}

# filter(inv_date < '2022-06-15')
# filter(inv_dat == lubridate::today())
# trees |> filter(parks_to_keep == 1)

park_trees <- trees |> #st_drop_geometry() |> # drops spatial information
    filter(dbhheight == '4.5ft') |>
    filter(department == 'Parks' & growspace == 'Yard/Park/Open space')

mapview(parks) + mapview(park_trees)

( # get species abundance and rank
  abundance_rank_parks <- trees |> 
    filter(!is.na(`Park Name`)) |> 
    st_drop_geometry() |> 
    group_by(Species = species_co) |> 
    summarise(Abundance = n()) |> 
    mutate(Rank = rank(-Abundance, ties.method = 'first')
           , plotname = ifelse(Rank < 11, Species, ''))
  )
  
(
  rac <- abundance_rank_parks |> 
    ggplot(aes(Rank, Abundance, label = plotname)) + 
    geom_point(size = 1) + 
    geom_line() + 
    ggrepel::geom_text_repel(max.overlaps = 50, size = 4, nudge_x=15) +
    labs(
      title = 'Rank Abundance Curve for Park Trees:'
      , subtitle = 'New Haven, CT'
    ) + 
    theme_bw(16) + 
    NULL
  )


# WORKS, is *GOOD*
# # RAC x (select) neighborhood
# ( # get species abundance and rank
#   abundance_rank_park <- trees |>
#     filter(!is.na(`Park Name`)) |> 
#     filter(parks_to_keep == 1) |> 
#     st_drop_geometry() |> 
#     # group_by(holc_grade, Species = species_co) |> # PUT SPECIES LAST
#     # group_by(neighborhood, Species = species_co) |> # PUT SPECIES LAST
#     group_by(`Park Name`, Species = species_co) |> # PUT SPECIES LAST
#     summarise(Abundance = n()) |> 
#     mutate(Rank = rank(-Abundance, ties.method = 'first')
#            , plotname = ifelse(Rank < 6, Species, ''))
#   )
# 
# # abundance_rank_park |> tail()
# 
# # graph out RAC by HOLC
# (
#   rac_select_parks <- abundance_rank_park |> 
#     ggplot(aes(Rank, Abundance, label = plotname)) + 
#     geom_line() + 
#     # geom_point(aes(color = holc_grade), size = 1) + #, shape = 21) +
#     # geom_point(aes(color = neighborhood), size = 1) + #, shape = 21) + 
#     geom_point(aes(color = `Park Name`), size = 2) + #, shape = 21) + 
#     ggrepel::geom_text_repel(max.overlaps = 100, size = 4, nudge_x = 15) +
#     scale_color_manual(values = holc_pal) + 
#     theme_bw(16) + 
#     theme(legend.position = 'none') + 
#     # facet_wrap(~holc_grade) + 
#     # facet_wrap(~neighborhood) + 
#     facet_wrap(~`Park Name`) + 
#     NULL
#   )


# RAC x parks (select vs all others)
( # get species abundance and rank
  abundance_rank_park_vs_all <- trees |>
    filter(!is.na(`Park Name`)) |> 
    mutate(park = ifelse(parks_to_keep == 1, `Park Name`, 'all other parks')) |> 
    st_drop_geometry() |> 
    # group_by(holc_grade, Species = species_co) |> # PUT SPECIES LAST
    # group_by(neighborhood, Species = species_co) |> # PUT SPECIES LAST
    group_by(park, Species = species_co) |> # PUT SPECIES LAST
    summarise(Abundance = n()) |> 
    mutate(Rank = rank(-Abundance, ties.method = 'first')
           , plotname = ifelse(Rank < 6, Species, ''))
  )

# abundance_rank_park |> tail()

# graph out RAC by HOLC
(
  rac_select_parks_all <- abundance_rank_park_vs_all |> 
    ggplot(aes(Rank, Abundance, label = plotname)) + 
    geom_line() + 
    geom_point(aes(color = park), size = 2) + #, shape = 21) + 
    ggrepel::geom_text_repel(max.overlaps = 100, size = 4, nudge_x = 15) +
    # scale_color_manual(values = holc_pal) + 
    theme_bw(16) + 
    theme(legend.position = 'none') + 
    facet_wrap(~park) + 
    NULL
  )




# RAC x parks (select vs all others)
( # get species abundance and rank
  abundance_rank_select_park <- trees |>
    filter(parks_to_keep == 1) |> 
    st_drop_geometry() |> 
    # group_by(holc_grade, Species = species_co) |> # PUT SPECIES LAST
    # group_by(neighborhood, Species = species_co) |> # PUT SPECIES LAST
    group_by(`Park Name`, Species = species_co) |> # PUT SPECIES LAST
    summarise(Abundance = n()) |> 
    mutate(Rank = rank(-Abundance, ties.method = 'first')
           , plotname = ifelse(Rank < 6, Species, ''))
  )

# abundance_rank_park |> tail()

# graph out RAC by HOLC
(
  rac_select_parks <- abundance_rank_select_park |> 
    ggplot(aes(Rank, Abundance, label = plotname)) + 
    geom_line() + 
    geom_point(aes(color = `Park Name`), size = 2) + #, shape = 21) + 
    ggrepel::geom_text_repel(max.overlaps = 100, size = 4, nudge_x = 15) +
    # scale_color_manual(values = holc_pal) + 
    theme_bw(16) + 
    theme(legend.position = 'none') + 
    facet_wrap(~`Park Name`) + 
    NULL
  )

# repeat with trees from THAT day in those parks vs the prviously collected trees

```

End here for MODs



## D street segments
```{r}

# TODO compare Bay Hanson's TreeCount to one generated with a spatial join of new Davey Resource Group trees
(
  st <- st_read('input_data/URI_Roads/URI_Roads_No_Highways_Jan2022.shp', as_tibble = TRUE) |> 
    st_transform(crs = st_crs(trees)) |> 
    dplyr::rename(TreeCount_Bay = TreeCount) %>%
    mutate(road_segment_length_m = as.numeric(st_length(.))) |> 
    rowid_to_column('street_id') |> 
    clean_names() |> 
    filter(road_segment_length_m > 21) # trim out small segments
 )

st |> mapview()

st |> glimpse()

st |> st_drop_geometry() |> select(road_segment_length_m) |> summary() # in METERS

st |> 
  ggplot(aes(road_segment_length_m)) + 
  # geom_density() + 
  geom_histogram(binwidth = 10) + 
  NULL

st |> 
  select(road_segment_length_m) |> 
  summary()


st |> filter(road_segment_length_m < 9) |> mapview() # TODO trim more!
st |> filter(road_segment_length_m < 22) |> mapview() # TODO trim more!
st |> arrange(road_segment_length_m) |> filter(road_segment_length_m < 22) |> mapview()
st |> filter(road_segment_length_m > 1200) |> mapview()

# unique ID?
dim(st)
n_distinct(st$street_id)
length(unique(st$fid_roadsf)) # aack!
length(unique(st$linearid))   # no

```




# 2 joins and tests
```{r}

# used for adding demographics from Neighborhood data
# tabular join to see if neighborhood code in trees and neighborhood file matches
trees |> left_join(neighs |> st_drop_geometry(), by = 'neighborhood') # Success
# trees |> anti_join(neighs |> st_drop_geometry(), by = 'neighborhood') |> mapview() # should be empty!, great

# join street trees to street segments spatially
# compare Bay's work.
tictoc::tic(); trees |> 
  # filter(department == 'Street Tree') |> # street trees only
  # sample_n(1e4) |> # used for testing
  st_join(st |> select(street_id, fullname, geoid, tree_count_bay)
          , left = TRUE
          , join = nngeo::st_nn, maxdist = 50, k = 1, progress = TRUE, parallel = 8) -> trees_w_st
tictoc::toc(); beepr::beep() # ~ 256s with just Street Trees, maxdist = 50 and parallel = 8

# # complete matches?
# trees |> filter(is.na(fullname)) |> mapview() + mapview(st, color = 'black')
# trees |> filter(is.na(street_id))|> mapview() + mapview(st, color = 'black')

# do the street names look right?
test |> st_drop_geometry() |> tabyl(street, fullname) # yes, yes they do.

trees |> st_drop_geometry() |> tabyl(nbrhood, neighborhood)

```

get top species per HOLC grade

# 3 community ecology metrics
## A RAC
```{r}

( # get species abundance and rank
  abundance_rank <- trees_w_st |> 
    st_drop_geometry() |> 
    group_by(Species = species_co) |> 
    summarise(Abundance = n()) |> 
    mutate(Rank = rank(-Abundance, ties.method = 'first')
           , plotname = ifelse(Rank < 11, Species, ''))
  )
  
(
  rac <- abundance_rank |> 
    ggplot(aes(Rank, Abundance, label = plotname)) + 
    geom_point(size = 1) + 
    geom_line() + 
    ggrepel::geom_text_repel(max.overlaps = 50, size=2.5, nudge_x=15) +
    theme_bw(16) + 
    NULL
  )


# RAC x HOLC
( # get species abundance and rank
  abundance_rank_holc <- trees_w_st |> 
    st_drop_geometry() |> 
    group_by(holc_grade, Species = species_co) |> # PUT SPECIES LAST
    summarise(Abundance = n()) |> 
    mutate(Rank = rank(-Abundance, ties.method = 'first')
           , plotname = ifelse(Rank < 6, Species, ''))
  )

# # 2x check
# abundance_rank_holc |> 
#   filter(holc_grade == 'A') |> 
#   arrange(Rank, Abundance)

# graph out RAC by HOLC
(
  rac_holc <- abundance_rank_holc |> 
    ggplot(aes(Rank, Abundance, label = plotname)) + 
    geom_line() + 
    geom_point(aes(color = holc_grade), size = 1) + #, shape = 21) + 
    ggrepel::geom_text_repel(max.overlaps = 100, size = 3, nudge_x = 15) +
    scale_color_manual(values = holc_pal) + 
    theme_bw(16) + 
    theme(legend.position = 'none') + 
    facet_wrap(~holc_grade) + 
    NULL
  )


# ggsave(file = paste0(getwd(), '/figures/XX_', Sys.Date(), '.png'),
#        height = 3*2, width = 6.5*2, dpi = 450)


```

## B species accumulation curves
```{r}

trees_w_st |> dim()
n_distinct(trees_w_st$id)

x <- trees_w_st |> 
  st_drop_geometry() |> 
  filter(!is.na(street_id)) |> 
  select(#id, 
    street_id, species_bo
    ) |>
  mutate(abundance = 1) |> 
  pivot_wider(  id_cols = street_id
              , names_from = 'species_bo'
              , values_from = abundance
              , values_fn = list(abundance = sum)
              , values_fill = 0) |> 
  clean_names() |> 
  column_to_rownames(var = 'street_id')


tictoc::tic(); out.all<- iNEXT(x, q=c(0,1,2),datatype="abundance"); tictoc::toc() # 164.696 sec elapsed
out.all$DataInfo # showing basic data information.
out.all$AsyEst # showing asymptotic diversity estimates.
out.all$iNextEst # showing diversity estimates with rarefied and extrapolated.

qlabels <- c("0" = "richness (q=0)", "1" = "Shannon's EFN (q=1)","2" = "Simpson's EFN (q=2)")

accum_alltree_all_q<-ggiNEXT(out.all, type=1, facet.var="order") + theme_bw(base_size=10)+ theme_bw(base_size=10)+ xlim(c(0,20000))+scale_colour_manual(values = holc_col,name="HOLC grade")+scale_fill_manual(values = holc_col, name="HOLC grade")+
  labs(x = "Number of individual trees sampled", y = "Tree species diversity", colour = "HOLC grade", shape = "HOLC grade", fill="HOLC grade")+
  theme(legend.position="bottom",legend.title=element_blank())+facet_wrap(~order, scales="free",labeller=labeller(order = qlabels))


tictoc::tic()
size_rare_estD <- iNEXT::estimateD(x, datatype = "abundance", base = "size", level = NULL,conf = 0.95); tictoc::toc()
print(size_rare_estD)
```


## sand box
```{r eval=FALSE, include=FALSE}

```



```{r, citations}
lapply(packages, citation)
```

Last knit on `r format(Sys.time())`

```{r eval=FALSE, include=FALSE}
system.time(save.image(file = paste0('saved_sessions/New_Haven_street_trees_redlining_'
                                     , gsub('[[:punct:]]', '-', Sys.time()), '.RData')))
```


# END

